---
title: "Population and T2D case"
author: "Sepi Sharif"
date: "9/27/2018"
output: html_document
---

```{r }
library(knitr)
library(dplyr)
library(narnia)
```
### "AllNewCleanData.txt" is the name of the master dataset made out of DM,RD,GOUT nadNPH datasets after filterring for NA, <18 years old, QCfaile and duplicated PATIENTID numbers.
### 940= T2DCase for the master dateset, 3408= T2DControl of the masterdataset
```{r }

AllNewCleanDataset<-read.delim( "../Data/AllNewCleanData.txt"  , header=TRUE, na.strings=c("NA",""," "))
AllNewCleanDataset%>% select(DIABETES) %>% table(., useNA='always') 
```
#### the name of the first column of the master dataset renamed to be matched with PCA datafile
```{r }
colnames(AllNewCleanDataset)[1]<-"PATIENTID"
```
### PCADataset
```{r }
ALLPCADATA<-read.delim( "../Data/PCAtestdata(1).txt"  , header=TRUE, na.strings=c("NA",""," "))
```
#### The masterdateset andPCA file merged to make the populations added to the datset
```{r }
AllNewCleanDatasetAndPCA<-merge(AllNewCleanDataset,ALLPCADATA,by="PATIENTID",all.x = T,all.y = F)

```
###Eastpolynesian population includind case and control after filterring for outliers 
### 381(T2DCase)-368 (T2DCase after filterring for outliers)=13 outliers case
###961 (T2DControl)- 947 (T2DControl after filterring for outliers)= 14 outliers
```{r }
EastPolynesiancasecontrol<-AllNewCleanDatasetAndPCA%>% filter(PCAETHBROAD=="East Polynesian")

EastPolynesiancasecontrol%>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 


#outliers for East polynesian continous dataset

vars2<-EastPolynesiancasecontrol[,c("AGECOL","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]

vars2nooutlier<-EastPolynesiancasecontrol[,c("PATIENTID","WAIST","WEIGHT","HEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]



j <- 0
for (i in vars2){
  j <- j+1
  mn <- mean(i, na.rm=T)
  ss <- sd(i, na.rm=T)
  outliers <- i > (mn + 4*ss) | i < (mn - 4*ss)
  table(outliers)
  var_range <- c(min(mn - 4*ss, min(i, na.rm=T)),
                 max(mn + 4*ss, max(i, na.rm=T)))
  hist(i, xlim=var_range, main=names(vars2)[j])
  abline(v = (mn + 4*ss))
  abline(v = (mn - 4*ss))
  abline(v =mn, col='blue')
  cat(names(vars2)[j], "\n")
  outliers %>%
    which() %>% 
    print()
  cat('\n')
}



EastPolynesiancasecontrolNoOutlier<-EastPolynesiancasecontrol[-c(1191,153,453,344,649,802,1244,
                                                                 127,923,1311,1055,309,631,905,946,963,1020,1049,1155,1231,1286,365,899,964,513,88,799),]


EastPolynesiancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 
```
###Westpolynesian population includind case and control after filterring for outliers 
### 200(T2DCase)-196 (T2DCase after filterring for outliers)=4 outliers case
###601 (T2DControl)- 589 (T2DControl after filterring for outliers)= 12 outliers
```{r }
WestPolynesiancasecontrol<-AllNewCleanDatasetAndPCA%>% filter(PCAETHBROAD =="West Polynesian")

WestPolynesiancasecontrol %>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 

#outliers for West polynesian continoud dataset

vars3<-WestPolynesiancasecontrol[,c("AGECOL","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]
vars3nooutlier<-WestPolynesiancasecontrol[,c("PATIENTID","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]


j <- 0
for (i in vars3){
  j <- j+1
  mn <- mean(i, na.rm=T)
  ss <- sd(i, na.rm=T)
  outliers <- i > (mn + 4*ss) | i < (mn - 4*ss)
  table(outliers)
  var_range <- c(min(mn - 4*ss, min(i, na.rm=T)),
                 max(mn + 4*ss, max(i, na.rm=T)))
  hist(i, xlim=var_range, main=names(vars3)[j])
  abline(v = (mn + 4*ss))
  abline(v = (mn - 4*ss))
  abline(v =mn, col='blue')
  cat(names(vars3)[j], "\n")
  outliers %>%
    which() %>% 
    print()
  cat('\n')
}


WestPolynesiancasecontrolNoOutlier<-WestPolynesiancasecontrol[-c(390,400,466,489,128,600,379,474,595,711,712,789,77,376,523,792),]


WestPolynesiancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 

```
###European population includind case and control after filterring for outliers 
### 263(T2DCase)-250 (T2DCase after filterring for outliers)=13 outliers case
###1441 (T2DControl)- 1410 (T2DControl after filterring for outliers)= 31 outliers
```{r }

Europeancasecontrol<-AllNewCleanDatasetAndPCA %>% filter(PCAETHBROAD=="European")

Europeancasecontrol %>% select (DIABETES) %>% table(., useNA='always')%>% cbind() 

#outliers for European population continous dataset

vars4<-Europeancasecontrol[,c("AGECOL","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]

vars4nooutlier<-Europeancasecontrol[,c("PATIENTID","WAIST","HEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]

j <- 0
for (i in vars4){
  j <- j+1
  mn <- mean(i, na.rm=T)
  ss <- sd(i, na.rm=T)
  outliers <- i > (mn + 4*ss) | i < (mn - 4*ss)
  table(outliers)
  var_range <- c(min(mn - 4*ss, min(i, na.rm=T)),
                 max(mn + 4*ss, max(i, na.rm=T)))
  hist(i, xlim=var_range, main=names(vars4)[j])
  abline(v = (mn + 4*ss))
  abline(v = (mn - 4*ss))
  abline(v =mn, col='blue')
  cat(names(vars4)[j], "\n")
  outliers %>%
    which() %>% 
    print()
  cat('\n')
}

#finding PATIENTID to remove outliers


EuropeancasecontrolNoOutlier<-Europeancasecontrol[-c(25,29,174,692,1511,544,780,1451,148,154,170,187,631,1169,1293,1609,1504,1525,1228,577,897,900,969,1047,1162,1187,1211,1278,1390,1417,1502,1553,1605,1679,532,933,1290,1590,1653,490,1047,1190,1502,1612,16224,1646,1704
                                                     ),]

EuropeancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()

```
###Mixedpolynesian population includind case and control after filterring for outliers 
### 27(T2DCase)-26 (T2DCase after filterring for outliers)=1 outliers case
###102 (T2DControl)- 98 (T2DControl after filterring for outliers)= 4 outliers
```{r }

Polynesiancasecontrol<-AllNewCleanDatasetAndPCA%>% filter(PCAETHBROAD =="Polynesian")

Polynesiancasecontrol %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()

#outliers for  polynesian continoud dataset

vars5<-Polynesiancasecontrol[,c("AGECOL","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]
vars5nooutlier<-Polynesiancasecontrol[,c("PATIENTID","WAIST","HEIGHT","WEIGHT","BMI","SYSTOLIC","AGECOL","DIASTOLIC","CHOLES","TRIGLY","HDL","LDL","COLHDL","SURICACID","EGFR")]

j <- 0
for (i in vars5){
  j <- j+1
  mn <- mean(i, na.rm=T)
  ss <- sd(i, na.rm=T)
  outliers <- i > (mn + 4*ss) | i < (mn - 4*ss)
  table(outliers)
  var_range <- c(min(mn - 4*ss, min(i, na.rm=T)),
                 max(mn + 4*ss, max(i, na.rm=T)))
  hist(i, xlim=var_range, main=names(vars5)[j])
  abline(v = (mn + 4*ss))
  abline(v = (mn - 4*ss))
  abline(v =mn, col='blue')
  cat(names(vars5)[j], "\n")
  outliers %>%
    which() %>% 
    print()
  cat('\n')
}

#finding PATIENTID to remove outliers


PolynesiancasecontrolNoOutlier<-Polynesiancasecontrol[-c(92,99,85,103,60),]



PolynesiancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()


```
##### n=329 is European and Non-polynesian ancestry for T2DCase, n= 1683 is European and Non-polynesian ancestry for T2DControl
```{r}

AllNewCleanDatasetAndPCA %>% select (PCAETHBROAD, DIABETES)  %>% table(., useNA='always') %>% cbind()

```
#### Different population T2D case and control with no duplications and outliers
```{r}
EastPolynesiancasecontrolNoOutlier%>% select(DIABETES) %>% table(., useNA='always') %>%cbind()
WestPolynesiancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()
EuropeancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always')%>% cbind()
PolynesiancasecontrolNoOutlier %>% select (DIABETES) %>% table(., useNA='always') %>% cbind()

```

