---
title: "Master dataset"
output: html_document
---
####the T2D case and control for the four datasets after , age<18, NA for T2D, QC filterring QC are as belwo. They have been used to make the master dataset: DMgenotypeQCindividualspassedQC,GOUTgenotypeQCindividualspassedQC,RDgenotypeQCindividualspassedQC,NPHgenotypeQCindividualspassedQC.

####T2DCase= 180, T2DControl= 58 for DMDatsetQCPassed which reads the DMgenotypeQCindividualspassedQC dataset (made after filterring for age<18, NA for T2D and QC) 
```{r }
library(dplyr)

DMDatasetQCPassed<-read.delim("../Data/DMgenotypeQCindividualspassedQC.txt" , header = T, na.strings = c(""," ","NA"))

DMDatasetQCPassed %>% select (DIABETES) %>% table()
```
### Making T2DCase and Control for DM dataset
```{r }
#T2DCase

DMCase<-DMDatasetQCPassed %>% select (PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                  HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                  CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter (DIABETES==2)



#T2DControl

DMControl<-DMDatasetQCPassed %>% select (PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                      HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                      CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter (DIABETES==1)

T2DDM<-rbind(DMCase, DMControl)

```
####3= borderline=maybe.  T2DCase= 584, T2DControl= 2970 for DMDatsetQCPassed which reads the GOUTgenotypeQCindividualspassedQC dataset (made after filterring for age<18, NA for T2D and QC
```{r }
GoutDatasetQCPassed<-read.delim("../Data/GOUTgenotypeQCindividualspassedQC.txt" , header = T, na.strings = c(""," ","NA"))
GoutDatasetQCPassed %>% select (DIABETES) %>% table()

```
### Making T2DCase and Control separately for GOUT dataset
```{r }
GOUTData<-read.delim("../Data/GOUTgenotypeQCindividualspassedQC.txt"  , header=TRUE, na.strings=c("NA",""," "))

GOUTCase<-GOUTData %>% select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                      HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GPGOUTAFFSTAT,GLUCMEASURED,
                                      CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR_SCL
) %>% filter (DIABETES==2)



#T2DControl

GOUTControl<-GOUTData %>% select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                              HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GPGOUTAFFSTAT,GLUCMEASURED,
                              CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR_SCL
) %>% filter (DIABETES==1)

T2DGOUT<-rbind(GOUTCase, GOUTControl)
```
#### T2DCase= 204, T2DControl= 146 for RrnalDatsetQCPassed which reads the RDenotypeQCindividualspassedQC dataset (made after filterring for age<18, NA for T2D and QC
```{r }
RenalDatasetQCPassed<-read.delim("../Data/RDgenotypeQCindividualspassedQC.txt" , header = T, na.strings = c(""," ","NA"))
RenalDatasetQCPassed %>% select (DIABETES) %>% table()

```
### Making T2DCase and Control separately for RD dataset
```{r }
#T2DCase

RenalData<-read.delim(  "../Data/RDgenotypeQCindividualspassedQC.txt"  , header=TRUE, na.strings=c("NA",""," "))


RenalDataCase<-RenalData %>% select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,INSULIN,
                                                     HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,HBA1C ,
                                                     CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter (DIABETES==2)



#T2DControl

RenalDataControl<-RenalData %>% select(PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,INSULIN,
                                    HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,HBA1C ,
                                    CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter (DIABETES==1)

T2DRD<-rbind(RenalDataCase, RenalDataControl)


```
#### T2DCase= 72, T2DControl= 234 for NPHDatsetQCPassed which reads the NPHgenotypeQCindividualspassedQC dataset (made after filterring for age<18, NA for T2D and QC
```{r }
NPHDatasetQCPassed<-read.delim("../Data/NPHgenotypeQCindividualspassedQC.txt" , header = T, na.strings = c(""," ","NA"))
NPHDatasetQCPassed %>% select (DIABETES) %>% table()
```
### Making T2DCase and Control separately for NPH dataset
```{r }
#T2DCase

NPHData<-read.delim( "../Data/NPHgenotypeQCindividualspassedQC.txt" , header=TRUE, na.strings=c("NA",""," "))


NPHDataCase<-NPHData%>% select (PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter (DIABETES==2)



#T2DControl
NPHDataControl<-NPHData%>% select (PATIENT,AGECOL,SEX,DIABETES,WEIGHT,HEIGHT,DIABETESTREAT,
                                HIBP,HEART,KIDNEY,WAIST,BMI,SYSTOLIC,DIASTOLIC,GOUTAFFSTAT,DIABINCULDEB,
                                CHOLES,TRIGLY,HDL,LDL,COLHDL,SURICACID,EGFR
) %>% filter (DIABETES==1)


T2DNPH<-rbind(NPHDataCase, NPHDataControl)

```
### Renaming the columns to be the same for rbinding
```{r }
##### making a masterdataset

colnames(T2DDM)

colnames(T2DGOUT)

colnames(T2DRD)

colnames(T2DNPH)



colnames(T2DGOUT)[16] <- "DIABINCULDEB"

colnames(T2DGOUT)[23] <- "EGFR"

colnames(T2DGOUT)[15]<-"GOUTAFFSTAT"


colnames(T2DNPH)[15]<-"GOUTAFFSTAT"

colnames(T2DRD)[7]<-"DIABETESTREAT" 
colnames(T2DRD)[16]<-"DIABINCULDEB"
colnames(T2DRD)[15]<-"GOUTAFFSTAT"


```
### rbinding the datasets and making the masterdatset, 1040= T2DCase with duplications, 3408=
###T2DControl with duplications
```{r }
RenalGOUTDMDNPHataset<-rbind(T2DDM,T2DGOUT,T2DRD,T2DNPH)

write.table (RenalGOUTDMDNPHataset,"../Data/AllNewCleanDatawithduplications.txt",sep="\t", row.names=F, quote = F)

RenalGOUTDMDNPHataset%>% select(DIABETES) %>% table(., useNA='always') %>% cbind()
```
####removing duplicated IDS(repeatedIDs), the output of this rbinding and removing duplications called "AllNewCleanData.txt". T2ControlduplicatedID= 3408-3369=39, T2DCaseduplicatedID= 1040-940=100
```{r }
RenalGOUTDMDNPHatasetOrdered<-RenalGOUTDMDNPHataset[order(RenalGOUTDMDNPHataset$DIABETES),]

RenalGOUTDMDNPHatasetNoDuplication<-RenalGOUTDMDNPHatasetOrdered[!duplicated(RenalGOUTDMDNPHatasetOrdered$PATIENT),]


RenalGOUTDMDNPHatasetNoDuplication%>% select(DIABETES) %>% table(., useNA='always') %>% cbind()
```
####RenalGOUTDMDNPHatasetNoDuplication exporting to Desktop and used as the "AllNewCleanData.txt" file for detecting ancestry populations with PCA.
```{r }
write.table (RenalGOUTDMDNPHatasetNoDuplication,"../Data/AllNewCleanData.txt",sep="\t", row.names=F, quote = F)
```
